### 2021-09-09 paper/localpsf.tex
writing proof for optimization problem

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = v_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $v_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We solve \eqref{eq:wi_optimization_problem_generic} efficiently via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item We form $f$ as follows:
\begin{equation*}
v(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
\end{equation*}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are computed by solving the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{v} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{v} \in \mathbb{R}^r$ is the vector with $i$th component $v_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation*}
S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
\end{equation*}
\end{enumerate}
In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.

Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, and these Poisson PDE solves can be performed cheaply with multigrid. All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to compute functions $\psi_i$ and $\theta_i$ associated with the new points. 
%, rather than $2(r+1)$ new PDE solves.

@@

@@

		\For{$i=1,2, \dots, r$}{
%			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.

			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_i$.

@@

			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_i$.

@@

		Form the $r \times r$ matrix $S$ with entries
		\begin{equation}
		\label{eq:defn_of_S_alg}
		S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.
		\end{equation}

@@

		\For{$i=1,2,\dots,r$}{
			Compute
			\begin{equation}
			\label{eq:definition_of_alpha_alg}
			\beta = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
			\end{equation}
			where $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $e_i = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is unit vector with $i^\text{th}$ component equal to one.

			Compute 
			\begin{equation}
			\label{eq:definition_of_mu_alg}
			\lambda = -S^{-1} \left(e_i - \beta \mathbf{1} \right).
			\end{equation}

			Form the weighting function $w_i$ given by
			\begin{equation}
			\label{eq:definition_of_wi_alg}
				w_i(x) = \beta - \sum_{k=1}^r \lambda_k \theta_k(x).
			\end{equation}
		}
	}
	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
	\label{alg:weighting_functions_incremental}
\end{algorithm2e}

\begin{thm}[Correctness of Algorithm \ref{alg:weighting_functions_incremental}]
\label{thm:wi_alg_is_correct}
	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
\end{thm}

>>>>

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = v_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $v_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
\begin{equation}
\label{eq:optimal_v}
v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
\end{equation}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{v} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{v} \in \mathbb{R}^r$ is the vector with $i$th component $v_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation}
\label{eq:S_matrix}
S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
\end{equation}
\end{enumerate}
In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.

Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, then post-processing these functions. This is shown in Algorithm \ref{alg:weighting_functions_incremental}. The Poisson PDE solves, while the most expensive part of this procedure, can be performed cheaply with multigrid.

%All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to compute functions $\psi_i$ and $\theta_i$ associated with the new points. 
%, rather than $2(r+1)$ new PDE solves.

@@

\begin{thm}[Solution of optimization problem \ref{eq:wi_optimization_problem_generic}]
	\label{thm:wi_alg_is_correct}
	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \ref{eq:wi_optimization_problem_generic}.
\end{thm}

\begin{proof}
	Let
	\begin{equation*}
		X := \{v \in H^2_N(\Omega) : v(x_i) = v_i \text{ for }i=1,\dots,r\}
	\end{equation*}
	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
	\begin{equation*}
	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,r\}
	\end{equation*}	
	denote the tangent space to $X$. Note that pointwise evaluation of functions in $H^2_N(\Omega)$ is well-defined because the Sobolev impedding theorem implies that $H^2_N(\Omega)$ is continuously imbedded in the spacde of continuous functions $C^0(\Omega)$ whenever the spatial dimension satisfies $d < 4$, and here we consider $d \in \{1,2,3\}$. Hence the optimization problem, the feasible set $X$ and the tangent space $X_0$ are well-defined.
\end{proof}

@@

		\For{$k=1,2, \dots, r$}{
%			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.

			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_k$.

@@

			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_k$.

@@

		Form the $r \times r$ matrix $S$ given in \eqref{eq:S_matrix}.

@@

		\For{$i=1,2,\dots,r$}{
			Solve the linear system
			\begin{equation*}
			\begin{bmatrix}
			S & \mathbf{1} \\ \mathbf{1}^T & 0
			\end{bmatrix}
			\begin{bmatrix}
			\mathbf{c}^{(i)} \\ \beta^{(i)}
			\end{bmatrix}
			=
			\begin{bmatrix}
			\mathbf{e}^{(i)} \\ 0
			\end{bmatrix}.
			\end{equation*}
			for $\mathbf{c}^{(i)}=(c_1^{(i)}, c_2^{(i)}, \dots, c_r^{(i)})$ and $\beta^{(i)}$. Here $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $\mathbf{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.

			Form the weighting function $w_i$ given by
			\begin{equation*}
			w_i(x) = \beta^{(i)} + \sum_{k=1}^r c_k^{(i)} \theta_k(x).
			\end{equation*}

		}
	}
	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
	\label{alg:weighting_functions_incremental}
\end{algorithm2e}
